A statistical mechanics approach to the simulation of extreme fluctuations of the time-averaged drag acting on an obstacle immersed in a turbulent flow

### 2018-12-26 draft_p2.tex
Presque fini la sous section sur l'AMS. Reste à écrire un pragraphe de perspectives

\author{Thibault Lestang}
\email{thibault.lestang@ens-lyon.fr}
\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
\author{Freddy Bouchet}
\email{freddy.bouchet@ens-lyon.fr}
\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
\author{Emmanuel Lévêque}
\email{emmanuel.leveque@ec-lyon.fr}
\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
\maketitle
\section{Introduction}

\section{Description of the test case for extreme drag fluctuations sampling}
\label{sec:test_flow}

@@

\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
\label{sec:ams}
The \acl{ams}~\cite{amspaper} inherits from the base ideas of \textit{splitting} algorithms~\cite{kahnharris}: the sampling of a rare event can be made easier by splitting the event
into a sequence of events that can be sampled with a higher probabilbity.
Even though a large body of variants have been proposed since the early 50s, the \ac{ams} appears as the most advanced formulation of multilevel splitting to date.
In recent years, it was applied to several problems

In the present paper we use a variant of the \ac{ams} algorithm called the \ac{tams}.
The algorithm is sketched in figure~\ref{figure_AMS}.
For further details about the algorithm and assoiated mathematical results, see~\cite{lestang2018} and references therein.
In the remainder of this section, we illustrate that the application of the \ac{TAMS}, fails to easily and efficiently sample rare drag fluctuations occuring in the test flow described in section~\ref{sec:test_flow}.
We analyse those difficulties in light of the conclusions of sections~\ref{sec:instantaneous_drag} and~\ref{sec:time_avg} on the dynamics of extreme drags.

To begin with, figure~\ref{fig:comparaison_temps_retour} illustrate expected results from the \ac{tams}, based on a one-dimensional \ac{ou} process.
In this case, it is clear that the algorithm allows for the sampling of very rare fluctuations for a computational cost that is many order of magnitude lower than their return time.

We then apply the \ac{tams} to the test flow, to sample extreme drag fluctuations on the square obstacle.

  \centering
  \includegraphics[width=.7\linewidth]{ams/figure_AMS}
  \caption{Illustration of one selection-mutation step in the \ac{ams} algorithm for the computation of the probability that an observable $A:\mathbb{R}^d\to \mathbb{R}$  reaches values larger than $Q$ over a trajectory of duration $T_a$.}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{ams/comparaison_temps_de_retour}
    \caption{Illustration of the efficiency of the \ac{tams} algorithm with respect to direct sampling. The amplitude of fluctuations for the \ac{ou} process~\eqref{eq:OUprocess_chap_ams} are represented along the Y-axis. Along the X-axis is represented the typical computational cost required to sample fluctuations of the corresponding amplitude in both the \ac{tams} and direct sampling. In the latter, the typical computational cost is simply the return time $r(a)$. The computational cost associated to a level $a$ is the \ac{tams} is denoted by $C(a)$. The orange line represents the evolution of the level $\mathcal{Q}_j^{\star}$ as a function of the cost $C(a)$. The blue stars represent the maximum of $x(t)$ over the resampled trajectory \emph{after} it is resampled, that is at iteration $j+1$. Finally, the solid black line is the analytical solution for the return time of amplitude $a$~\cite{lestang_computing_2018}.}
  \label{fig:comparaison_temps_de_retour}
\end{figure}

>>>>

\author{Thibault Lestang}
\email{thibault.lestang@ens-lyon.fr}
\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
\author{Freddy Bouchet}
\email{freddy.bouchet@ens-lyon.fr}
\affiliation{Univ Lyon, ENS de Lyon, Univ Claude Bernard, CNRS, Laboratoire de Physique, F-69342 Lyon, France}
\author{Emmanuel Lévêque}
\email{emmanuel.leveque@ec-lyon.fr}
\affiliation{Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
\maketitle
\section{Introduction}
The purpose of this paper is twofold.
\section{Description of the test case for extreme drag fluctuations sampling}
\label{sec:test_flow}

@@

\subsection{Extreme instantaneous drag forces with the \acl{ams} algorithm}
\label{sec:ams}
% Introduction to AMS
The \acl{ams}~\cite{amspaper} inherits from the base ideas of \textit{splitting} algorithms~\cite{kahnharris}: the sampling of a rare event is made easier by splitting the event into a sequence of events that can be sampled with a higher probabilbity.
Even though a large body of variants have been proposed since the early 50s, the \ac{ams} appears as the most advanced formulation of multilevel splitting to date.
In recent years, it was applied to several problems, ranging from the computation of the dissociation time of biomolecules~\cite{teo2016} to the simulation of rare relaminarisations in a stochastic model of wall turbulence~\cite{rolland2017}.

% Quick description of TAMS
In the present paper we use a variant of the \ac{ams} algorithm called the \ac{tams}.
The \ac{ams}, and therefore the \ac{tams}, relies on the definition of a \textit{score function} $\xi (\mathbf{x}(t))$.
It is an observable of the dynamics that acts as a measure of the distance of the current state $\mathbf{x}(t)$ to the region of the state space that correspond to the rare event(s) of interest.
The \ac{ams} algorithm is an iterative procedure, that starts from an ensemble of $N$ independant trajectories of the dynamics over a time-interval $[0;T_a]$.
At each iteration, trajectories with the lowest maxima of the score function over $[0;T_a]$ are discarded and replaced by trajectories that achieved a higher value of the score function at some point of their
history,
The algorithm is sketched in figure~\ref{figure_AMS}.
For further details about the algorithm and assoiated mathematical results, see~\cite{lestang2018} and references therein.

% Example: OU process
We first describe the result of the \ac{tams} on a simple case.
We consider a one-dimensional \ac{ou} process
\begin{equation}
  \label{eq:ou}
  \dot{x} = -x + \sqrt{2\epsilon}\eta (t)  
\end{equation}

and apply the \ac{tams} to simulate trajectories $\{\mathbf{x}(t)\}_{0\leq t \leq T_a}$ with $\underset{0\leq t \leq T_a}{\max} x(t) \geq a$ with $a$ very large with respect to typical fluctuations of $x(t)$.
Figure~\ref{fig:compraison_temps_de_retour} illustrate the amplitude of the maximum over trajectories along the iterations of the algorithm.
The computational cost of the algorithm after iteration $j$ is the computational cost of simulating the $N$ initial trajectories, as well as the resampling of the $j$ trajectories.
Figure~\ref{fig:compraison_temps_de_retour} illustrates that the \ac{tams} samples extreme fluctuations $x \geq a$ for a computational cost several order of magnitude lower than the return time of such fluctuations.
Undoubtedly, an \ac{ou} process is an oversimplified dynamics to showcase the efficiency of the \ac{ams} and \ac{tams}.
Indeed, the state space is one-dimensional and the choice of the score function poses no questions: it is simply the observable $x$ itself.
In addition, the noise term in~\eqref{eq:ou} has no correlations in time, which means that newly generated trajectories will quickly separate from their ancesotrs.

% Application to turbulent flow past an obstacle
Can we obtain similar results by applying the \ac{tams} to the test flow described in\ref{sec:test_flow} ?n
The objective is to sample trajectories that display extreme fluctuatons of the drag $f_d$ acting on the square obstacle. However, by contrast with a simple \ac{ou} process, the phase space in higly-dimensinal and the choice of the score function $xi (\mathbf{x}(t))$ is not so clear.
In the following we use the \ac{tams} with $\xi \equiv f_d$, which is the most simple choice for the score function.
For such complex dynamics however, there is no resons for the drag itself to be a good measure of how far is a given state in phase space to regions where the drag is extreme.
Unfortunately, the choice of better score functions rely on a deep understanding of the physics of the rare events of interest, whcih is precisely what we are tryping to achieve.

% Commentaire figure instant_256_20

Figure~\ref{fig:instant_256_20} displays similar information as figure~\ref{fig:comparaison_temps_de_retour} for $N=256$ initial trajectories over $T_a = 20 \tau_c$.
Additionally, figure~\ref{fig:instant_256_20} shows the distribuion of the maximum of the drag among the initial trajectories.
Over the first iterations of the algorithm, trajectories with the lowest maximum of the score function are discarded, and new trajectories with higher maximum are resampled.
These maxima are depicted by the stars in figure~\ref{fig:instant_256_20}.
However, the maximum over the resampled trajectories never exceeds the amplitude of the highest maximum in the initial set of trajectories.

% Commentaire de la figure iter_181

Figure~\ref{fig:instant_256_20} can be better understood by visualizing the set of the last resampled trajectories. 
One can see that, ultimately, trajectories have been resampled from the same initial trajectory.
We explain this phenonemon as follows.
Because the flow dynamics hae a memory, it takes a certain time before a resampled trajectory separate from its parent.
This is clearly visible on figure~\ref{fig:instant_256_20}.
In our case, this memory originates from the deterministic nature of the dynamics, which implies that two trajectories starting from infinitesimally close starting point will overlap over a timescale called
the <<Lyapunov timesacle>>~\cite{lyapunovRef}.
However, we observe that the duration of extreme drag fluctuations is typically 5 times smaller that the Lyapunov timescale of the drag proces.
This means that the resampling of a trajectory based on the value of the drag at $t^{\star}$ cannot lead to larger drag values for $t^{\star} \leq t \leq \tau_L$.
For $t > \tau_L$, the drag process has no memory of the drag fluctuations on which the resampling was based and the probability of observing an extreme fluctuation over $\tau_l \leq t \leq T_a$ is very low.
The difference between the typical duration of drag fluctuations $\tau_c$ and the Lyapunov timescale $\tau_0$ is linked to the convective nature of the flow.
Indeed, fluid structures responsible for a particular value of the drag on the obstacle at a time $t^{\star}$ are swept away by the mean flow over a timescale $\tau_c$.

% perspectives

\begin{itemize}
\item more trajectories
\item different score functino?
\item Change algorithm
\end{itemize}
\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{ams/figure_AMS}
  \caption{Illustration of one selection-mutation step in the \ac{ams} algorithm for the computation of the probability that an observable $A:\mathbb{R}^d\to \mathbb{R}$  reaches values larger than $Q$ over a trajectory of duration $T_a$.}
\end{figure}
\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{ams/comparaison_temps_de_retour}
    \caption{Illustration of the efficiency of the \ac{tams} algorithm with respect to direct sampling. The amplitude of fluctuations for the \ac{ou} process~\eqref{eq:OUprocess_chap_ams} are represented along the Y-axis. Along the X-axis is represented the typical computational cost required to sample fluctuations of the corresponding amplitude in both the \ac{tams} and direct sampling. In the latter, the typical computational cost is simply the return time $r(a)$. The computational cost associated to a level $a$ is the \ac{tams} is denoted by $C(a)$. The orange line represents the evolution of the level $\mathcal{Q}_j^{\star}$ as a function of the cost $C(a)$. The blue stars represent the maximum of $x(t)$ over the resampled trajectory \emph{after} it is resampled, that is at iteration $j+1$. Finally, the solid black line is the analytical solution for the return time of amplitude $a$~\cite{lestang_computing_2018}.}
  \label{fig:comparaison_temps_de_retour}
\end{figure}
